---
title: "Assessing the impact of motion on connectivity"
author: "Tim Rittman"
date: "23/06/2015"
output: pdf_document
---

This script is designed to assess the correlation between spike percentage and connection strength - whether it exists and what the correct cut-off for spike percentage should be in the GENFI data.

```{r, echo=FALSE, results='asis'}
library(plyr)
library(ggplot2)
library(xtable)
library(xlsx)

genfiDir = "/home/tim/GENFI/GENFI_camgrid_20150525/"
setwd(genfiDir)

inFile = "d2_degree_wt_local"

# import data
dF = read.table(inFile, header = TRUE, na.strings = "NA")

# convert diagnostic label to a factor and rename
dF$GS = as.factor(dF$GS)
dF$GS = revalue(dF$GS, c("0" = "gene negative", "1"="gene positive", "2"="affected"))


nodeNames = names(dF)[sapply(names(dF), function(x) grepl("X",x))]

# Re
dF.stacked = stack(dF[,nodeNames])
dF.stacked = data.frame(dF.stacked, GS=dF$GS, gene=dF$gene, wbic=dF$wbic)

dF.wb = ddply(dF.stacked, .(gene, wbic, GS), summarise,
              values = mean(values, na.rm = TRUE)
              )

# import file with spike percentage data
sp <- read.xlsx("../all_sm_thld10_SP.xlsx", sheetIndex = 1)
sp.sub <- data.frame(wbic=sp$id, spMean=sp$mean)
dF.sp <- merge(dF.wb, sp.sub, by="wbic")

```

We'll plot the correlation value between the connection strength and spike percentage for various thresholds of spike percentage.

The first plot shows the variation with spike percentage threshold of the Pearson correlation coefficient between connection strength and spike percentage.

The second plot shows the p-value of this correlation as the spike percentage threshold varies. The dashed line represents a significant value of 0.05.

```{r, echo=FALSE, results='asis'}
getCorr <- function(n,x){
  ct <- cor.test(x[x$spMean<n,"spMean"], x[x$spMean<n,"values"])
  return(list(ct[["statistic"]][[1]],
              ct[["p.value"]][[1]])
         )
}

cutOffs <- seq(2,40, by=1)
ctVals <- sapply(cutOffs, getCorr, x=dF.sp)
dF.ctVals <- data.frame(matrix(unlist(ctVals), byrow = TRUE, ncol = length(ctVals[,1])))
names(dF.ctVals) <- c("r", "p")

# plot the results
dF.plot <- data.frame(cutOffs=cutOffs, dF.ctVals)

p <- ggplot(dF.plot, aes_string(x="cutOffs", y="r"))
p <- p + geom_line()
p <- p + theme_bw()
p <- p + labs(y="correlation coefficient", x="spike percentage threshold")
p

p <- ggplot(dF.plot, aes_string(x="cutOffs", y="p"))
p <- p + geom_line(colour="red")
p <- p + geom_hline(y=0.05, linetype="dashed", colour="black")
p <- p + theme_bw()
p <- p + labs(y="p-value", x="spike percentage threshold")
p

```